Calculating Biological Module Enrichment or Depletion and Visualizing Data on Large-scale Molecular Maps with the R packages \pkg{ACSNMineR} and \pkg{RNaviCell}

### 2016-01-15 jss/article.tex
[skip ci] small corrections

Malignant gliomas, the most common subtype of primary brain tumors, are
aggressive, highly invasive, and neurologically destructive tumors considered to
be among the deadliest of human cancers. In its most aggressive manifestation,
glioblastoma (GBM), median survival ranges from 9 to 12 months, despite maximum
treatment efforts \citep{maher2001malignant}. In this study we have analyzed
whole-genome mutation data generated by the TCGA project on hundreds of
patients. More specifically, we parsed the MAF (Mutation Annotation Format) GBM
files produced by different sequencing centers to count and calculate gene
mutation frequencies. We kept the mutations having a status likely to disturb
the target protein's function (i.e, Frame\_Shift\_Del, Frame\_Shift\_Ins,
In\_Frame\_Del, In\_Frame\_Ins, Missense\_Mutation, Nonsense\_Mutation). In
total, we collected mutations for xxx genes in a total of xxx samples. In order
to retain the most frequently mutated genes, we calculated frequencies across
all mutated samples, and kept genes having a frequency greater than 1\%. We
further labelled genes having a frequency greater than 1\% and less than 5\% as "1"
and genes highly mutated (frequency higher than 5\%) as "2".

>>>>

Malignant gliomas, the most common subtype of primary brain tumors, are
aggressive, highly invasive, and neurologically destructive tumors considered
to be among the deadliest of human cancers. In its most aggressive
manifestation, glioblastoma (GBM), median survival ranges from 9 to 12 months,
despite maximum treatment efforts \citep{maher2001malignant}. In this study we
have analyzed whole-genome mutation data generated by the TCGA project on
hundreds of patients. More specifically, we parsed the MAF (Mutation Annotation
Format) GBM files produced by different sequencing centers to count and
calculate gene mutation frequencies. We kept the mutations having a status
likely to disturb the target protein's function (i.e, Frame\_Shift\_Del,
Nonstop\_Mutation, In\_Frame\_Del, In\_Frame\_Ins, Missense\_Mutation,
Nonsense\_Mutation, Splice\_Site, Translation\_Start\_Site). In total, we
collected mutations for more than 13,000 genes in a total of 379 mutated
samples. In order to retain the most frequently mutated genes, we calculated
frequencies across all mutated samples, and kept genes having a frequency
greater than 1\% (3,293 genes). We further labelled genes having a frequency
greater than 1\% and less than 5\% as "1" and genes highly mutated (frequency
higher than 5\%) as "2".
